---
title: "College QB Weekly Stats Report"
subtitle: "Power 4 Quarterback Rankings by EPA/Play"
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    code-fold: true
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
params:
  season: 2025
  min_attempts: 50
  conferences: ["Big Ten", "SEC", "Big 12", "ACC"]
---

```{r setup}
library(cfbfastR)
library(dplyr)
library(reactable)
library(reactablefmtr)
library(htmltools)
```

```{r fetch-data}
#| cache: true

# Fetch play-by-play data for the season
pbp_data <- cfbfastR::load_cfb_pbp(seasons = params$season)

# Calculate QB stats from play-by-play
qb_stats <- pbp_data |>
  filter(
    pass == 1,
    !is.na(passer_player_name),
    passer_player_name != ""
  ) |>
  group_by(
    Player = passer_player_name,
    Team = pos_team,
    Conference = offense_conference
  ) |>
  summarize(
    Games = n_distinct(week),
    ATT = n(),
    CMP = sum(completion, na.rm = TRUE),
    YDS = sum(yards_gained, na.rm = TRUE),
    TD = sum(pass_td, na.rm = TRUE),
    INT = sum(int, na.rm = TRUE),
    `Total EPA` = round(sum(EPA, na.rm = TRUE), 1),
    .groups = "drop"
  ) |>
  mutate(
    `CMP%` = round(CMP / ATT * 100, 1),
    `Y/A` = round(YDS / ATT, 1),
    `EPA/Play` = round(`Total EPA` / ATT, 3)
  ) |>
  filter(
    ATT >= params$min_attempts,
    Conference %in% params$conferences
  ) |>
  arrange(desc(`EPA/Play`)) |>
  select(Player, Team, Conference, Games, ATT, CMP, `CMP%`, YDS, `Y/A`, TD, INT, `Total EPA`, `EPA/Play`)
```

## Overview

This report ranks **`r nrow(qb_stats)`** quarterbacks from the **`r params$season`** season with at least **`r params$min_attempts`** pass attempts in selected conferences.

::: {.callout-note}
## What is EPA?
**Expected Points Added (EPA)** measures how much each play changes the expected points for the offense. **EPA/Play** is the average EPA per pass attempt - the gold standard for evaluating QB efficiency.
:::

## Quarterback Rankings

Use the search box and column filters to explore the data. Click column headers to sort.

```{r table}
reactable(
  qb_stats,
  searchable = TRUE,
  filterable = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  striped = TRUE,
  compact = TRUE,
  defaultPageSize = 25,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(10, 25, 50, 100),
  defaultSorted = list(`EPA/Play` = "desc"),
  theme = reactableTheme(
    borderColor = "#dfe2e5",
    stripedColor = "#f6f8fa",
    highlightColor = "#f0f5ff",
    cellPadding = "8px 12px",
    style = list(fontFamily = "-apple-system, BlinkMacSystemFont, Segoe UI, Helvetica, Arial, sans-serif")
  ),
  columns = list(
    Player = colDef(
      minWidth = 150,
      sticky = "left",
      style = list(fontWeight = "bold")
    ),
    Team = colDef(minWidth = 120),
    Conference = colDef(minWidth = 100),
    Games = colDef(minWidth = 60, align = "center"),
    ATT = colDef(minWidth = 60, align = "center"),
    CMP = colDef(minWidth = 60, align = "center"),
    `CMP%` = colDef(minWidth = 70, align = "center", format = colFormat(suffix = "%")),
    YDS = colDef(minWidth = 70, align = "center", format = colFormat(separators = TRUE)),
    `Y/A` = colDef(minWidth = 60, align = "center"),
    TD = colDef(
      minWidth = 50,
      align = "center",
      style = list(color = "#28a745", fontWeight = "bold")
    ),
    INT = colDef(
      minWidth = 50,
      align = "center",
      style = list(color = "#dc3545")
    ),
    `Total EPA` = colDef(
      minWidth = 90,
      align = "center",
      style = function(value) {
        color <- if (value > 0) "#28a745" else if (value < 0) "#dc3545" else "#6c757d"
        list(color = color, fontWeight = "bold")
      }
    ),
    `EPA/Play` = colDef(
      minWidth = 90,
      align = "center",
      style = function(value) {
        color <- if (value > 0) "#28a745" else if (value < 0) "#dc3545" else "#6c757d"
        list(color = color, fontWeight = "bold")
      }
    )
  )
)
```

## Top 10 by EPA/Play

```{r top10}
top_10 <- head(qb_stats, 10)

reactable(
  top_10,
  bordered = TRUE,
  striped = TRUE,
  highlight = TRUE,
  theme = reactableTheme(
    borderColor = "#dfe2e5",
    stripedColor = "#f0f5ff"
  ),
  columns = list(
    Player = colDef(minWidth = 150, style = list(fontWeight = "bold")),
    Team = colDef(minWidth = 120),
    Conference = colDef(minWidth = 100),
    Games = colDef(minWidth = 50, align = "center"),
    ATT = colDef(minWidth = 50, align = "center"),
    CMP = colDef(minWidth = 50, align = "center"),
    `CMP%` = colDef(minWidth = 60, align = "center"),
    YDS = colDef(minWidth = 60, align = "center"),
    `Y/A` = colDef(minWidth = 50, align = "center"),
    TD = colDef(minWidth = 40, align = "center", style = list(color = "#28a745")),
    INT = colDef(minWidth = 40, align = "center", style = list(color = "#dc3545")),
    `Total EPA` = colDef(minWidth = 80, align = "center"),
    `EPA/Play` = colDef(minWidth = 80, align = "center", style = list(fontWeight = "bold"))
  )
)
```

---

*Data sourced from [cfbfastR](https://cfbfastR.sportsdataverse.org/) | Report generated: `r Sys.Date()`*
